Fitting the Undetected: Light-curve Feature Estimation in the Presence of Non-detections

### 2012-04-11 tex/ms.tex
adding graphical model
commented graphical model b/c wouldn't compile
changed prior on b_i to normal

% Import the natbib package for bibliography inclusion. To get it to work right, first typeset the bibliography (BibTex), Command-Shift-B.
% Then, typeset the document with LaTeX, Command-Shift-L or just Command-T.
% To get proper reference citations using aastex (preprint manuscripts), first typeset as emulateapj with BibTex to create the BibTex 
% files and then typeset again as aastex to get the manuscript with references.
\usepackage{natbib}
\usepackage{epsfig,graphicx,color}
\usepackage{longtable}

@@

We model the observed flux, $\fobs$, as a Gaussian distribution with variance that has both measurement ($\sigma^2_i$) and model ($s^2_{\mu}$) contributions.  In the case that a detection is made, we require that the observed brightness be greater than the brightness of the detection limit ($\fobs = \ftrue \ge b_i$) while in the case that no detection is made, we require that the brightness (if it could be measured) be less than the detection limit ($\ftrue < b_i$).  To derive the likelihood of the observed $\fobs$ given $\sigma^2_i$, $\theta$ and $I$, we must integrate over the prior distribution of the unknown $b_i$,
\begin{eqnarray}\displaystyle
p(\fobs |\sigma^2_i,\theta,I) &=& \left\{\begin{array}{ll}
  N(\ftrue | \mu_i,  \sigma^2_i + s^2_{\mu})\,  \int_0^{\ftrue} p(b_i | \theta)\, \dd b_i & \mbox{if $\fobs \ne \texttt{NA}$} \\
  \int_{0}^{\infty} \int_{0}^{b_i} N(\ftrue | \mu_i, \sigma^2_i + s^2_{\mu})\, p(b_i | \theta)\, \dd \ftrue\, \dd b_i & \mbox{if $\fobs = \texttt{NA}$} \\
\end{array}\right.\label{eq:mlik}
\\
p(b_i|\theta) &=& N(b_i|B,V_B)
\label{eq:bprior}
\\
\theta &\equiv& (\omega, A_0, \{A_k, B_k\}_{k=1}^K, s^2_\mu, B, V_B, \cdots) \quad ,
\end{eqnarray}
where we have introduced the hyperparameters $B$ and $V_B$ for the Gaussian prior distribution of $b_i$.  In Equation \ref{eq:mlik}, in the epochs for which a detection was made ($\fobs \ne \texttt{NA}$), we marginalize over the unknown $b_i$ from $0$ (low brightness limit) to $\ftrue$, enforcing that the detection limit be fainter than the observed brightness.  Likewise, in the epochs for which no detection was made ($\fobs = \texttt{NA}$), we integrate the joint $(\ftrue, b_i)$ likelihood over all possible values of $b_i$ and over the unknown $\ftrue$ from $0$ to $b_i$, ensuring that the brightness (if it were able to be observed) be fainter than the detection limit.

In the above, we have assumed no extra information on each of the $b_i$ values besides the knowledge of whether a detection was made on that epoch.  Hence, we draw, in Equation \ref{eq:bprior}, each $b_i$ value from a global prior distribution which is the same at all epochs.  If instead, we are given an estimate of $b_i$ plus its error distribution for each epoch (which, in principle can be inferred from the raw telescope images), we can replace Equation \ref{eq:bprior} with a different distribution per epoch.  In the case that the $b_i$ are assumed to be completely known (without error), the data likelihood of $\fobs$ becomes
\begin{eqnarray}\displaystyle
p(\fobs |\sigma^2_i,\theta,I, b_i) &=& \left\{\begin{array}{ll}
  N(\fobs | \mu_i,  \sigma^2_i + s^2_{\mu})\,  I(\fobs \ge b_i) & \mbox{if $\fobs \ne \texttt{NA}$} \\
 \int_{0}^{b_i} N(\ftrue | \mu_i, \sigma^2_i + s^2_{\mu})\, \dd \ftrue & \mbox{if $\fobs = \texttt{NA}$} \\
\end{array}\right.\label{eq:mlik_s}
\quad ,
\end{eqnarray}
where the boolean indictor function, $I(\ftrue \ge b_i)$, is 1 if $\ftrue \ge b_i$ and 0 otherwise.

@@

>>>>

% Import the natbib package for bibliography inclusion. To get it to work right, first typeset the bibliography (BibTex), Command-Shift-B.
% Then, typeset the document with LaTeX, Command-Shift-L or just Command-T.
% To get proper reference citations using aastex (preprint manuscripts), first typeset as emulateapj with BibTex to create the BibTex
% files and then typeset again as aastex to get the manuscript with references.
\usepackage{natbib}
\usepackage{epsfig,graphicx,color}
\usepackage{longtable}
\usepackage{amsmath}

@@

We model the observed flux, $\fobs$, as a Gaussian distribution with variance that has both measurement ($\sigma^2_i$) and model ($s^2_{\mu}$) contributions.  In the case that a detection is made, we require that the observed brightness be greater than the brightness of the detection limit ($\fobs = \ftrue \ge b_i$) while in the case that no detection is made, we require that the brightness (if it could be measured) be less than the detection limit ($\ftrue < b_i$).  To derive the likelihood of the observed $\fobs$ given $\sigma^2_i$, $\theta$ and $I$, we must integrate over the prior distribution of the unknown $b_i$,
\begin{eqnarray}\displaystyle
p(\fobs |\sigma^2_i,\theta,I) &=& \left\{\begin{array}{ll}
  N(\ftrue | \mu_i,  \sigma^2_i + s^2_{\mu})\,  \int_{-\infty}^{\ftrue} p(b_i | \theta)\, \dd b_i & \mbox{if $\fobs \ne \texttt{NA}$} \\
  \int_{-\infty}^{\infty} \int_{-\infty}^{b_i} N(\ftrue | \mu_i, \sigma^2_i + s^2_{\mu})\, p(b_i | \theta)\, \dd \ftrue\, \dd b_i & \mbox{if $\fobs = \texttt{NA}$} \\
\end{array}\right.\label{eq:mlik}
\\
p(b_i|\theta) &=& N(b_i|B,V_B)
\label{eq:bprior}
\\
\theta &\equiv& (\omega, A_0, \{A_k, B_k\}_{k=1}^K, s^2_\mu, B, V_B, \cdots) \quad ,
\end{eqnarray}
where we have introduced the hyperparameters $B$ and $V_B$ for the Gaussian prior distribution of $b_i$.  In Equation \ref{eq:mlik}, in the epochs for which a detection was made ($\fobs \ne \texttt{NA}$), we marginalize over the unknown $b_i$ from $0$ (low brightness limit) to $\ftrue$, enforcing that the detection limit be fainter than the observed brightness.  In the epochs for which no detection was made ($\fobs = \texttt{NA}$), we compute the probability that the unknown detection threshold $b_i$ is greater than the unknown true flux $f_i$. When both $b_i$ and $f_i$ are normally distributed, this can be computed using a single call to a Gaussian cumulative distribution function, specifically
\begin{equation*}
\mathbb{P}(b_i \geq f_i) = \int_{-\infty}^0 N(x|B-\mu_i,V_B + \sigma_i^2 + s_{\mu}^2) dx
\end{equation*}

In the above, we have assumed no extra information on each of the $b_i$ values besides the knowledge of whether a detection was made on that epoch.  Hence, we draw, in Equation \ref{eq:bprior}, each $b_i$ value from a global prior distribution which is the same at all epochs.  If instead, we are given an estimate of $b_i$ plus its error distribution for each epoch (which, in principle can be inferred from the raw telescope images), we can replace Equation \ref{eq:bprior} with a different distribution per epoch.  In the case that the $b_i$ are assumed to be completely known (without error), the data likelihood of $\fobs$ becomes
\begin{eqnarray}\displaystyle
p(\fobs |\sigma^2_i,\theta,I, b_i) &=& \left\{\begin{array}{ll}
  N(\fobs | \mu_i,  \sigma^2_i + s^2_{\mu})\,  I(\fobs \ge b_i) & \mbox{if $\fobs \ne \texttt{NA}$} \\
 \int_{-\infty}^{b_i} N(\ftrue | \mu_i, \sigma^2_i + s^2_{\mu})\, \dd \ftrue & \mbox{if $\fobs = \texttt{NA}$} \\
\end{array}\right.\label{eq:mlik_s}
\quad ,
\end{eqnarray}
where the boolean indictor function, $I(\ftrue \ge b_i)$, is 1 if $\ftrue \ge b_i$ and 0 otherwise.

@@

\begin{figure}[hbtp]
  \begin{center}
    \includegraphics[width=\textwidth,%
      trim=3cm 12cm 5cm 3cm, clip=true]{gm/gm.pdf}
  \end{center}
  \caption{The graphical representation of the probability distribution.
    \label{fig:graph}}
\end{figure}
